%% ----------------------------------------------------------------
%% thesis.tex
%% ---------------------------------------------------------------- 
\documentclass[a4paper]{18monthreport}

\usepackage[super,sort&compress]{natbib}            % Use Natbib style for the refs.
\usepackage{subfiles}

\hypersetup{colorlinks=true}   % Set to false for black/white printing
\graphicspath{{'C:\Users\soob\Documents\EnergyStoragePhD\phd_thesis_template-master\graphics'}}    % Location of your graphics files

\renewcommand{\bibname}{References}

%% ----------------------------------------------------------------
\begin{document}
\frontmatter
\title      {Interplay between Economic and Environmental Objectives for Renewable Energy Co-Located Storage Systems}
\date       {March 2018}
\author     {Susan Isaya Sun}
\supervisor {Dr. Richard G. A. Wills; Dr. Andrew J. Chipperfield}
\university {University of Southampton}
\department {Energy Technology}
\group      {Energy Technology Group}
\faculty    {Faculty of Engineering and the Environment}

\maketitle

\begin{abstract}
Abstract goes here.
\end{abstract}

\tableofcontents
\listoffigures
\listoftables
\lstlistoflistings
\listofsymbols{llc}{
$a$ & description & unit\\
$b$ & description & unit\\
$c$ & description & unit}
\listofacronyms{ll}{
ACR & acronym\\
ACR & acronym\\
ACR & acronym}
\declaration{Sun et al. 2017}
\acknowledgements{Thanks to no one.}
\mainmatter
%% ----------------------------------------------------------------
\chapter{Introduction}
\label{sec:Introduction}
\subfile{18month_intro}

\chapter{The Value Proposition}
\label{sec:The Value Proposition}
\subfile{18month_valueprop}

\chapter{Predicting the Price of Second-Life Electric Vehicle Batteries}
\label{sec:Predicting the Price of Second-Life Electric Vehicle Batteries}
\subfile{18month_2ndlife}

\chapter{Projections of Grid Emissions Intensities}
\label{sec:Projections of Grid Emissions Intensities}
\subfile{18month_gridemissions}

\chapter{Modelling Battery Degradation}
\label{sec:Modelling Battery Degradation}
\subfile{18month_battdeg}

\chapter{Calculating Connection Charges for Alternative Systems}
The text for another chapter goes here.

\chapter{Calculation Methodology}
The text for another chapter goes here.

\chapter{Sizing the System}
The text for another chapter goes here.

\chapter{Discussion}
The text for another chapter goes here.

\chapter{Conclusions}
The text for another chapter goes here.


\appendix{}
The text for an appendix goes here.

\appendix{}
The text for an appendix goes here.

%% ----------------------------------------------------------------
\backmatter
\bibliography{thesisrefs}{}
\bibliographystyle{plain}
\end{document}
%% ----------------------------------------------------------------
